---
title: "example"
output: html_document
---

### RMD Set-Up

The first chunk is where knitr options go, and the second line is the most critical. Typically RMDs run out of their own directory, but because I have data and code separated, that can make navigation complicated. This normalizePath just resets the RMD path to the main directory for the repository where the R Project is, so we can navigate more easily to data and code.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../")) # Again, don't skip this line - because the R Project working directory is set as "...GitHub/project-design/" but this .RMD is under "...GitHub/project-design/RMS/", this sets the document working directory backwards one step to be the same as the R Project file ("...GitHub/project-design/"). This way I don't have to put "../" in front of every single file or path call
```


Once the pathing is set, we can run the config.R file to initialize the Outputs subdirectories, set up iteration versioning, and load packages. Look inside this file for more information on how this is setup.

NOTE: Because I use date-based version folders, if you want to do the same, make sure you change the date here **before** running source.
```{r config, echo = FALSE}
### Set the date in YYYY-MM-DD format:
date <- "2025-06-17"
source("R/config.R")
```

With that done, you should now have all folders created and available to put data in.

Now I'll go into some examples for accessing data and creating output figures, but from here on out it's basically the same as working in a normal R project.

First make a demo dataset

```{r make_df, echo = FALSE}
species <- c("A", "B", "C", "D")
count <- c("14", "36", "2", "28")
demo <- data.frame(Species = species, Count = count)
```

Make a bar plot of the count data and write the image to our figure output folder
```{r plot, echo = FALSE}
ggplot(demo, aes(x = Species, y = Count)) + 
  geom_bar(stat = "identity")

# Save this plot to an object
plot <- ggplot(demo, aes(x = Species, y = Count)) + 
  geom_bar(stat = "identity")

# Now save the plot to our figures folder
ggsave("plot.png", plot, path = path_fig, width = 6, height = 4)
```

Some R packages output objects that don't fit neatly into a ggsave or write.csv function, but I still want to save them as intermediary backups or for workflow verification, and saveRDS() is useful for that.
To make saveRDS work with relative pathing, it's easiest to first pass the name and output path as a vector, and then use that vector as an input.

```{r}
ExportPath <- paste(path_out, "RDS_export.rds", sep = "/") # pastes our output path and a name for our RDS object with / as a separator
ExportPath # Now this appears as a complete path + file name

saveRDS(demo, ExportPath) # We'll use the demo df for the test, even though it could also be exported with write.csv()

# To load an RDS, you would need to create the file path and name again, then call the output folder
ExportPath <- paste(path_out, "RDS_export.rds", sep = "/") # Pretend like we didn't just do this 
ExportPath
demo2 <- readRDS(ExportPath, path_out)
```


